\bf Visual Diagnsotics of SAO Models

### 2017-02-11 paper.Rnw
more intro/lit review material

\noindent%
{\it Keywords:}  Visual inference, Lineup protocol, \hh{XXX Other keywords?}.
\vfill

@@

The guiding principles of model visualization are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that could benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out from behind-the-scenes and into the light of visualization will provide researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which are attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we can gain a better understanding of these correlations and find ways to deal with them and rectify their effects.

@@

@@

\section{Visualizations}
\hh{The visualizations should address most elements of structure described before, i.e. follow along in the setup (for descriptive and teaching purposes), the fitting (understanding the model and interpreting the results) and then diagnostics (how well does the model fit the data - where are the most differences to the actual data/what are the sensitive parameters?) }

>>>>

\noindent%
{\it Keywords:} social network analysis, model visualization \hh{Other keywords?}
\vfill

@@

The guiding principles of model visualization are (1) display the model in the data space, (2) collections of models are more informative than singletons, and (3) explore how algorithms work, instead of just looking at the final result. Stochastic actor-oriented models are a prime example of a set of models that could benefit greatly from application of model visualization. First, the models themselves include a continuous-time Markov chain (CTMC) that is completely hidden in the model fitting process. Bringing the CTMC out from behind-the-scenes and into the light of visualization will provide researchers with insights into the underlying features of the model. Secondly, SAOMs allow for a great deal of parameters that can be added to the objective function, each of which are attached to a network statistic. These statistics are often somewhat, if not highly, correlated, and so are the associated parameters. By visualizing collections of SAOMs, we can gain a better understanding of these correlations and find ways to deal with them and rectify their effects. Finally, estimation of the parameters in a SAOM relies on a Robins-Monro algorithm, and convergence checks rely on simulations from the fitted model. Again, in these steps are largely hidden from the researcher computing them with software. We hope that by bringing these background elements and others into the foreground, stochastic actor-oriented models will be better understood and more accessible to social network analysts.

@@

The stochastic actor-oriented models are very aptly named: they model dynamic networks that can account for the individual node characteristics when making the edge changes. The object of analysis is a social network that has been observed at several discrete time points. The nodes in the network can have associated covariate variables that may be incorporated into modeling the tie changes in the network. The model itself has two primary pieces: the rate of change of ties, and the objective function of the nodes, which determines the exact change that gets made.

We start with a set of network observations at $M$ discrete time points: $x(t_1), x(t_2) \dots, x(t_M)$. The first observation, $x(t_1)$ is always conditioned on. The number of nodes / actors / vertices is constant in all time points and is denoted $n$.  The tie variables are directed and  binary, and are denoted at $x_{ij}(t_m)$ for time point $m$, and this is sometimes abbreviated to $x_{ij}$ in notation. If there is a tie from node $i$ to node $j$, then $x_{ij}=1$, and $x_{ij}=0$ otherwise.

@@

\section{Model Visualizations}

\subsection{View the model in the data space}
\subsection{Collections are more informative than singletons}
\subsection{Explore algorithms, not just end result}
\hh{The visualizations should address most elements of structure described before, i.e. follow along in the setup (for descriptive and teaching purposes), the fitting (understanding the model and interpreting the results) and then diagnostics (how well does the model fit the data - where are the most differences to the actual data/what are the sensitive parameters?) }
